---
title: "Mini-Project 4: Text Analysis"
format:  
  html:
    code-fold: false
---

This mini-project explores sentiment, uniqueness, and trends accross Tupac's songs. The data obtained from <https://www.kaggle.com/datasets/leopoldofacci/tupac-dataset-lyrics-analyse>.

The data originally contained: "Lyrics", "Song", "Artist". I manually added the album names and will add years for time trajectory analysis in the final project.

```{r}
#Import relevant libraries
library(dplyr)
library(widyr)
library(tidyverse)
library(stringr)
library(tidytext)
library(DataEditR)
library(wordcloud2)
library(wordcloud)
library(RColorBrewer)
library(tidyr)
library(tidytext)
```

```{r}
#Read the data, and create a dictionary with regular expressions of cuss-words to filter out later.
tupac_untidy <- read.csv("~/Sds 264 F24/lyrics-TUPAC.csv")
cuss <- read.csv("~/Sds 264 F24/cuss2.csv")
cuss_words_pattern <- paste0("\\b(", paste(cuss$word, collapse = "|"), ")\\b") 
```

```{r}
#Create an unnested version of the dataset using str functions and regular expressions
tupac_sep <- tupac_untidy |> 
  unnest_tokens(word, Lyrics) |> 
  anti_join(cuss, by = "word")

tupac_tidy <- tupac_untidy |> 
  mutate(Lyrics = str_replace_all(Lyrics, "\\n", " ")) 



```

```{r}
#Through regular expressions, find out what are the largest numbers said in a song in the dataset
tupac_sep |> 
  mutate(number = as.numeric(str_extract(word, "\\b\\d+\\b"))) |>  
  filter(!is.na(number) & Song != "When I Get Free II") |>  
  slice_max(number, n = 10, with_ties = FALSE) |> 
  select(Song, number)

  

```
>> Here we can see that there are a lot of numbers, with some research we can find out that in the song "Outlawz" 2pac says "Things are not the same since 1995". 1995 is the year when 2pac was released pending an appeal of a conviction in 1995

>>In Soulja's Story, 1993 refers to a line "Straight soldier, 1993 onward", I am unsure what it means as the song was released in 1991.

```{r}
#Find words that are repeating twice in a row using regular expressions
tupac_tidy |> 
  mutate(repeated_words = str_extract_all(Lyrics, "\\b(\\w+)\\b[\\s]+\\1\\b")) |> 
  unnest(repeated_words) |> 
  filter(!is.na(repeated_words)) |> 
  distinct(Song, repeated_words) |> 
  select(Song, repeated_words)
```
>> Some of the twice repeating words are silly and some are serious, for example "La la" and "murder murder". This shows the duality of 2pac.

```{r}
#Through str functions, find out longest words:
tupac_sep |> 
  distinct(word, .keep_all = TRUE) |>  # Keep only distinct words
  mutate(word_length = str_length(word)) |>  # Calculate the length of each word
  filter(Song!="All Eyez On Me") |> 
  slice_max(word_length, n = 10, with_ties = FALSE)  # Select the top 10 longest words without ties

```

>>What could be the longest words in old-school rap? We see a famous catchphrase of 2pac which is yaknowhati'msayin	appearing in different songs, other words include: responsibilities, institutionalized, multimillionare, penitentiary, and congratulation. If anything- we can sense repeating themes of worry, money, and lifestyle.

```{r}
#Analyze the sentiment of positivity vs negativity in songs across albums
bing_sentiment <- get_sentiments("bing")

sentiment_scores <- tupac_sep |> 
  filter(!is.na(word)) |> 
  inner_join(bing_sentiment, by = "word") |>  #join sentiments and lyrics
  group_by(Album, sentiment) |> 
  summarize(sentiment_count = n(), .groups = 'drop') |> 
  ungroup()

ggplot(sentiment_scores, aes(x = Album, y = sentiment_count, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Positive and Negative Sentiment of Tupac Albums",
    x = "Album",
    y = "Sentiment Count",
    fill = "Sentiment"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_manual(values = c("positive" = "blue", "negative" = "black")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))

print(sentiment_scores)


 
```
>> Abnove, we can see the album with the highest positivity and negativity is All Eyez On Me. This album contains versatile songs and lyrics and as such it is not a surprise to see that it has the highest count of both sentiments.

```{r}
#Find the fear sentiment in 2pac songs and create a wordcloud
nrc_sentiment <- get_sentiments("nrc")

tupac_nrc <- tupac_sep|> 
  filter(!is.na(word))|> 
  inner_join(nrc_sentiment, by = "word")|> 
  filter(sentiment == "fear")|>   # Focus only on fear sentiment
  count(Album, word, sort = TRUE)|>  
  distinct(word, .keep_all = TRUE)|>   # Remove duplicates per album
  ungroup()

top_fear_words <- tupac_nrc|> 
  group_by(Album)|> 
  slice_max(n, n = 7)|>   
  ungroup()

wordcloud(
  words = top_fear_words$word, 
  freq = top_fear_words$n, 
  max.words = 200,  
  random.order = FALSE,  
  rot.per = 0.2,  
  scale = c(4, 0.5), 
  colors = brewer.pal(8, "Set1")  
)

```
>> The wordcloud above highlights words with a fearing sentiment to them. Some words include "hate, cemetery, wicked, and god as the highest fear sentiment.

```{r}
#A tf-idf plot with bigrams per album
tupac_ngram <- tupac_tidy |>
  unnest_tokens(word, Lyrics) |> 
  anti_join(stop_words, by = "word") |>  # Remove stop words
  anti_join(cuss, by = "word") |>  # Remove cuss words 
  group_by(Song, Album) |> 
  summarise(Lyrics = str_c(word, collapse = " "), .groups = "drop") |>  # Collapse words into single Lyrics column
  unnest_tokens(bigram, Lyrics, token = "ngrams", n = 2) |> 
  filter(!is.na(bigram))  # Filter out NA bigrams

# Calculating TF-IDF and visualizing
bigram_tf_idf <- tupac_ngram |> 
  count(Album, bigram, sort = TRUE) |> 
  bind_tf_idf(bigram, Album, n)

bigram_tf_idf |>
  group_by(Album) |> 
  arrange(desc(tf_idf)) |> 
  slice_max(tf_idf, n = 10, with_ties = FALSE) |>  # Select top 10 bigrams per Album
  ungroup() |>  
  ggplot(aes(x = fct_reorder(bigram, tf_idf), y = tf_idf, fill = Album)) +  # Reorder bigrams based on tf_idf
    geom_col(show.legend = FALSE) +  # Create bar plot
    coord_flip() +  # Flip coordinates for horizontal bars
    facet_wrap(~Album, scales = "free")

```

>> According to the td-idf statistic and analysis we can see a lot of 2pac's uniqueness in lyrics comes from themes of identity and even location (as California appears multpile times). For example, in 2Pacalypse now, 2Pac highlights his racial identity in sensitive detail. While "Greatest Hits" includes more lifestyle songs.

#In conclusion, Tupac's music has a lot of versatality- with a wide variety of emotions and topics. 
